---
title: "Reproducible Research: Peer Assessment 1"
output: 
  html_document:
    keep_md: true
---


## Loading and preprocessing the data
```{r}
data<-read.csv(unz("activity.zip","activity.csv"),stringsAsFactors=FALSE)
#data<-read.csv(unz("activity.zip","activity.csv"))
data$interval<-as.character(data$interval)


data$interval<-sapply(X=data$interval,function(time)
  
{
  
  if (nchar(time)==1)
    paste("00",":0",time,sep="")
  else if (nchar(time)==2)
    paste("00:",time,sep="")
  else if (nchar(time)==3)
    paste("0",substr(time,1,1),":",substr(time,2,3),sep="")
  else if(nchar(time)==4)
    paste(substr(time,1,2),":",substr(time,3,4),sep="")
  
  
}
)


library(lubridate)

data$newdate<-paste(data$date,data$interval,sep=" ")

#data$newdate<-as.Date(strptime(data$newdate,format="%Y-%m-%d %H:%M"))

processeddata<-data[complete.cases(data),]
```
## What is mean total number of steps taken per day?
```{r}
avgsteps<-tapply(processeddata$steps,processeddata$date,FUN=sum)
print(avgsteps)
```

## histogram of the total number of steps taken each day
```{r}
hist(avgsteps)
```

## Calculate and report the mean and median of the total number of steps taken per day
```{r}
meanavgsteps<-tapply(processeddata$steps,processeddata$date,FUN=mean)
print(meanavgsteps)

medianavgsteps<-tapply(processeddata$steps,processeddata$date,FUN=median)
print(medianavgsteps)

```


## What is the average daily activity pattern?
```{r}
avgsteps1<-tapply(processeddata$steps,processeddata$interval,FUN=mean)
library(ggplot2)

dfavgsteps<-data.frame(levels(factor(processeddata$interval)),avgsteps1)
colnames(dfavgsteps)<-c("interval","steps")
test$interval<-strptime(dfavgsteps$interval,format="%H:%M")
library(scales)
p<-ggplot(dfavgsteps,aes(x=interval,y=steps,group=1))+geom_line()+scale_x_datetime(labels = date_format("%H:%M"))

print(p)

```
## Imputing missing values

## Calculate and report the total number of missing values in the dataset (i.e. the total number of rows with NAs)
```{r}
print(nrow(data)-nrow(processeddata))
```

## filling in all of the missing values in the dataset using the mean for that particular interval
```{r}
newdata<-data

for(i  in 1:nrow(data))
{
  
  if (is.na(data$steps[i])) {newdata$steps[i]<-mean(data$steps[data$interval==data$interval[i]],na.rm=TRUE)}
}
    
summary(newdata)


```
## histogram of the total number of steps taken each day and Calculate and report the mean and median total number of steps taken per day
```{r}
avgsteps<-tapply(newdata$steps,newdata$date,FUN=sum)
print(avgsteps)
hist(avgsteps)

meanavgsteps<-tapply(newdata$steps,newdata$date,FUN=mean)
print(meanavgsteps)

medianavgsteps<-tapply(newdata$steps,newdata$date,FUN=median)
print(medianavgsteps)

```

## Are there differences in activity patterns between weekdays and weekends?
```{r}
newdata$isweekend=sapply(as.Date(newdata$date),function(x){
  if(weekdays(x)=="Sunday" || weekdays(x)=="Saturday")
    "weekend"
  else
    "weekday"  
})
levels(newdata$isweekend)<-factor(newdata$isweekend)
newdata$interval<-strptime(newdata$interval,format="%H:%M")
library(plyr)
mm <- ddply(newdata, c("isweekend","interval"), summarise, mmpg = mean(steps))
newplot<-ggplot(mm, aes(x = interval,y=mmpg)) +geom_line()+scale_x_datetime(labels = date_format("%H:%M"))+facet_grid(.~isweekend)
print(newplot)
```